Energy and the CES production function

### 2015-10-19 Articles/Article_EcolEcon_2015/EcolEcon_2015A.Rnw
Split models chunk into smaller chunks.

<<Models>>=
fileNameCDCES <- "EcolEcon2015ModelsCDCES.rds"
fileNameCSP <- "EcolEcon2015ModelsCSP.rds"
fileNameExp <- "EcolEcon2015ModelsExp.rds"
if (!rerunModels){
  # load from the file on disk
  models <- readRDS(fileNameCDCES)
  modelsCSP <- readRDS(fileNameCSP)
  modelsExp <- readRDS(fileNameExp)
} else {   ### PLEASE USE COMMENTS TO MARK WHERE LONG CHUNKS END; BETTER: AVOID LONG CHUNKS, WHERE POSSIBLE

  ### This has not been inspected by rjp

  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.

  #
  # Models that are exponential only, y = A * exp(lambda*t)
  #
  modelsExp <- list()
  alphaExp <- 0
  betaExp <- 0
  gammaExp <- 0
  for (country in countries){
    # Replace with better code later. 
    # Would like to say cdModel(formula, data, factorShares=c(alpha_1 = 0, alpha_2 = 0, alpha_3 = 0))
    # or something like that.
    form <- keep_summands(formulas[[country]][["unadjusted"]], n = c(-1, -2, -3)) # Keep none of the factors of production
    yVar <- as.name(summands(form, n=1, right=FALSE))
    tVar <- as.name(summands(form, n=1))
    data <- historicalData[[country]]
    expMod <- lm(log(eval(yVar)) ~ eval(tVar), data=data)
    attr(expMod, "naturalCoeffs") <- data.frame(lambda=coef(expMod)[[2]],
                                                  logscale=coef(expMod)[[1]], 
                                                  scale=exp(coef(expMod)[[1]]),
                                                  alpha_1=alphaExp,
                                                  alpha_2=betaExp, 
                                                  alpha_3=gammaExp)
    expMod$formula <- form
    sdata <- subset(data, select=all.vars(form))
    sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
    expMod$data <- sdata
    expMod$response <- sdata$iGDP 
    modelsExp[[country]] <- expMod
  }
  # Save exponential models to disk.
  saveRDS(modelsExp, file=fileNameExp)

  #
  # Models that adhere to the cost-share theorem
  #

  modelsCSP <- list()
  # Output elasticities based on cost share theorem
  alphaCSP <- 0.3 # Fraction of national accounts spent on capital
  betaCSP <- 0.7 # Fraction of national accounts spent on labor
  for (country in countries){
    for (adj in c("unadjusted", "adjusted")){
      # Replace with better code later. 
      # Would like to say cdModel(formula, data, factorShares=c(alpha_1 = 0.3, alpha_2 = 0.7, alpha_3 = 0))
      # or something like that.
      form <- keep_summands(formulas[[country]][[adj]], n=-3) # Remove energy
      yVar <- as.name(summands(form, n=1, right=FALSE))
      kVar <- as.name(summands(form, n=1))
      lVar <- as.name(summands(form, n=2))
      tVar <- as.name(summands(form, n=3))
      data <- historicalData[[country]]
      # Cobb-Douglas
      cdModCSP <- 
        lm(log(eval(yVar)) - alphaCSP*log(eval(kVar)) - betaCSP*log(eval(lVar)) ~ eval(tVar), 
           data = data)
      # cdModCSP is an "lm" object. It is not a "CDEmodel" object. 
      # Thus, it needs its own "naturalCoeffs" attribute if it is to work correctly in the "naturalCoef" function.
      attr(cdModCSP, "naturalCoeffs") <- data.frame(lambda=coef(cdModCSP)[[2]],
                                                    logscale=coef(cdModCSP)[[1]], 
                                                    scale=exp(coef(cdModCSP)[[1]]),
                                                    alpha_1=alphaCSP,
                                                    alpha_2=betaCSP, 
                                                    alpha_3=0)
      cdModCSP$formula <- form
      sdata <- subset(data, select=all.vars(form))
      sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
      cdModCSP$data <- sdata
      cdModCSP$response <- sdata$iGDP
      modelsCSP[[country]][[adj]][["noE"]][["CD"]] <- cdModCSP
      # CES
      # The only CES models that can adhere to the CSP are models where
      # alpha values are constant with respect to time. 
      # I.e., models where energy is not included AND where rho_1 = 0.
      # CES models where rho_1 = 0 are the Cobb-Douglas model.
      # Convienently, we previously fit that model.
      cesModCSP <- cdModCSP
      # Again, we need a customized "naturalCoeffs" attribute so that 
      # the "naturalCoef" function will do the right thing.
      # "naturalCoeffs" can be based on the same "naturalCoeffs"
      # from the Cobb-Douglas model, but a few additions are needed.
      attr(cesModCSP, "naturalCoeffs") <- 
        naturalCoef(cdModCSP) %>%
        mutate(
          delta_1 = naturalCoef(cdModCSP)[["alpha_1"]],
          delta = 1,
          rho_1 = 0, 
          rho = NA,
          sigma_1 = 1,
          sigma_12 = 1,
          sigma = NA,
          sigma_12_3 = NA
        )
      modelsCSP[[country]][[adj]][["noE"]][["CES"]][["kl"]] <- cesModCSP
    }
  }
  # Save cost-share theorem models to disk.
  saveRDS(modelsCSP, file=fileNameCSP)

  #
  # Models where all parameters are estimated (models don't adhere to the cost share theorem)
  #

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term (3rd term on RHS)
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2), digits = 10)
      nestsWithE <- list(kle=c(1,2,3), lek=c(2,3,1), ekl=c(3,1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- lapply(nestsWithE, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest, digits = 10)
        })
    }      
  }
  # Save models to disk.
  saveRDS(models, file=fileNameCDCES)
}  ### END ELSE FROM WAY UP THERE???
@

<<Models2>>=
### TIME FOR A NEW R CHUNK -- That helps detect the end of that long code chunk

### This function is the same in each model group below.  DRY (Don't repeat yourself).
### Using this makes it (a) easier to see what's going on, and (b) easier to make adjustments
### that affect all of the situations simultaneously.

### General hint:  If you find yourself wanting to copy and paste a chunk of code, there is usually
### a better way to do what you are about to do.

model2df <- function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  ### Don't name non-lists with variables whose names sound like a list.
  ### I'm guessing you just didn't know what kind of data you were working with here. 
  ### Either way, it's a bad thing.
  id_components <- strsplit(id, split = "[.]")[[1]]

@@

# Put all the cost-share theorem (CSP) model coefficients into a data frame
AllCoeffsCSP <- 
  leaf_apply(modelsCSP, class = c("lm"), f = model2df) %>% 
  bind_rows() %>%
  mutate(CSP = "w/ CSP")

@@

### It looks like this could also be done the otehr way arround:  We could scoop up all the models
### into AllCoeffs and then filter to the subsets.

# Get reference values for lambda and sse.
### This could be done without saving ref if we have no other need for it
### I've renamed ref to Ref since it is a data frame.
Ref <- 
  AllCoeffs %>% 
  filter(model=="exp") %>% 
  select(Country, sse, lambda) %>%
  rename(sse_ref = sse, lambda_ref = lambda)
AllCoeffs <- merge(AllCoeffs, Ref)

# Calculate differences relative to the reference model
### If all we need are the differences, we could skip the part where we put the intermediate stuff 
### into AllCoefs

AllCoeffs <- 
  AllCoeffs %>%
  mutate(
    Dlambda = abs(lambda) - abs(lambda_ref), 
    Dsse = sse - sse_ref,
    percDlambda = Dlambda / abs(lambda_ref) * 100, 
    percDsse = Dsse / sse_ref * 100
  )

@@

%' <<>>=
%' leaf_apply(models, function(mod, id) class(mod), class="cesModel")
%' @

@@

AllOE <- 
  leaf_apply(
    models, 
    class = "cesModel", 
    mod2oedf
  ) %>% 
  bind_rows()
@

@@

### One line for each operation:  first leaf_apply, then rbind, then mutate -- Clarity by Code Convention
temp_data_CSP <- 
  leaf_apply(modelsCSP, class=c("plm", "lm"), model2df) %>% 
  bind_rows() %>%
  mutate(CSP = "w/ CSP")

>>>>

<<Models_CSP>>=
fileNameCSP <- "EcolEcon2015ModelsCSP.rds"
if (!rerunModels){
  # load from the file on disk
  modelsCSP <- readRDS(fileNameCSP)
} else {
  #
  # Models that adhere to the cost-share theorem
  #

  modelsCSP <- list()
  # Output elasticities based on cost share principle
  alphaCSP <- 0.3 # Fraction of national accounts spent on capital
  betaCSP <- 0.7 # Fraction of national accounts spent on labor
  for (country in countries){
    for (adj in c("unadjusted", "adjusted")){
      # Replace with better code later. 
      # Would like to say cdModel(formula, data, factorShares=c(alpha_1 = 0.3, alpha_2 = 0.7, alpha_3 = 0))
      # or something like that.
      form <- keep_summands(formulas[[country]][[adj]], n=-3) # Remove energy
      yVar <- as.name(summands(form, n=1, right=FALSE))
      kVar <- as.name(summands(form, n=1))
      lVar <- as.name(summands(form, n=2))
      tVar <- as.name(summands(form, n=3))
      data <- historicalData[[country]]
      # Cobb-Douglas
      cdModCSP <- 
        lm(log(eval(yVar)) - alphaCSP*log(eval(kVar)) - betaCSP*log(eval(lVar)) ~ eval(tVar), 
           data = data)
      # cdModCSP is an "lm" object. It is not a "CDEmodel" object. 
      # Thus, it needs its own "naturalCoeffs" attribute if it is to work correctly in the "naturalCoef" function.
      attr(cdModCSP, "naturalCoeffs") <- data.frame(lambda=coef(cdModCSP)[[2]],
                                                    logscale=coef(cdModCSP)[[1]], 
                                                    scale=exp(coef(cdModCSP)[[1]]),
                                                    alpha_1=alphaCSP,
                                                    alpha_2=betaCSP, 
                                                    alpha_3=0
                                                    )
      cdModCSP$formula <- form
      sdata <- subset(data, select=all.vars(form))
      sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
      cdModCSP$data <- sdata
      cdModCSP$response <- sdata$iGDP
      modelsCSP[[country]][[adj]][["noE"]][["CD"]] <- cdModCSP
      # CES
      # The only CES models that can adhere to the CSP are models where
      # alpha values are constant with respect to time. 
      # I.e., models where energy is not included AND where rho_1 = 0.
      # CES models where rho_1 = 0 are the Cobb-Douglas model.
      # Convienently, we previously fit that model.
      cesModCSP <- cdModCSP
      # Again, we need a customized "naturalCoeffs" attribute so that 
      # the "naturalCoef" function will do the right thing.
      # "naturalCoeffs" can be based on the same "naturalCoeffs"
      # from the Cobb-Douglas model, but a few additions are needed.
      attr(cesModCSP, "naturalCoeffs") <- 
        naturalCoef(cdModCSP) %>%
        mutate(
          delta_1 = naturalCoef(cdModCSP)[["alpha_1"]],
          delta = 1,
          rho_1 = 0, 
          rho = NA,
          sigma_1 = 1,
          sigma_12 = 1,
          sigma = NA,
          sigma_12_3 = NA
        )
      # Add output elasticities (as a function of time) to this model.
      # This would have been done by a "fortify" function, 
      # but this is not a typical CES model, so we can't "fortify" it.
      # Add other metadata, too.
      cesModCSP$data <- cesModCSP$data %>%
        mutate(
          oe1 = alphaCSP,
          oe2 = betaCSP,
          oe3 = 0,
          Country = country, 
          flavor  = ifelse(adj=="unadjusted", "Unadjusted", "Quality-adjusted"),
          energy  = "Without energy",
          model   = "CES", 
          nest    = "kl",
          CSP     = "w/ CSP"
        )
      modelsCSP[[country]][[adj]][["noE"]][["CES"]][["kl"]] <- cesModCSP
    }
  }
  # Save cost-share theorem models to disk.
  saveRDS(modelsCSP, file=fileNameCSP)
}

@

<<Models_Exp>>=
fileNameExp <- "EcolEcon2015ModelsExp.rds"
if (!rerunModels){
  # load from the file on disk
  modelsExp <- readRDS(fileNameExp)
} else {
  #
  # Models that are exponential only, y = A * exp(lambda*t)
  #
  modelsExp <- list()
  alphaExp <- 0
  betaExp <- 0
  gammaExp <- 0
  for (country in countries){
    # Replace with better code later. 
    # Would like to say cdModel(formula, data, factorShares=c(alpha_1 = 0, alpha_2 = 0, alpha_3 = 0))
    # or something like that.
    form <- keep_summands(formulas[[country]][["unadjusted"]], n = c(-1, -2, -3)) # Keep none of the factors of production
    yVar <- as.name(summands(form, n=1, right=FALSE))
    tVar <- as.name(summands(form, n=1))
    data <- historicalData[[country]]
    expMod <- lm(log(eval(yVar)) ~ eval(tVar), data=data)
    attr(expMod, "naturalCoeffs") <- data.frame(lambda=coef(expMod)[[2]],
                                                  logscale=coef(expMod)[[1]], 
                                                  scale=exp(coef(expMod)[[1]]),
                                                  alpha_1=alphaExp,
                                                  alpha_2=betaExp, 
                                                  alpha_3=gammaExp)
    expMod$formula <- form
    sdata <- subset(data, select=all.vars(form))
    sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
    expMod$data <- sdata
    expMod$response <- sdata$iGDP 
    modelsExp[[country]] <- expMod
  }
  # Save exponential models to disk.
  saveRDS(modelsExp, file=fileNameExp)
}

@

<<Models_CDCES>>=
fileNameCDCES <- "EcolEcon2015ModelsCDCES.rds"
if (!rerunModels){
  # load from the file on disk
  models <- readRDS(fileNameCDCES)
} else {
  #
  # Models where all parameters are estimated (models don't adhere to the cost share theorem)
  #
  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term (3rd term on RHS)
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2), digits = 10)
      nestsWithE <- list(kle=c(1,2,3), lek=c(2,3,1), ekl=c(3,1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- lapply(nestsWithE, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest, digits = 10)
        })
    }      
  }
  # Save models to disk.
  saveRDS(models, file=fileNameCDCES)  
}
@

<<AllCoeffs>>=
model2df <- function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  ### Don't name non-lists with variables whose names sound like a list.
  ### I'm guessing you just didn't know what kind of data you were working with here. 
  ### Either way, it's a bad thing.
  id_components <- strsplit(id, split = "[.]")[[1]]

@@

# Put all the cost-share principle (CSP) model coefficients into a data frame
AllCoeffsCSP <- 
  leaf_apply(modelsCSP, class = c("lm"), f = model2df) %>% 
  bind_rows() %>%
  mutate(CSP = "w/ CSP")

@@

@@

@@

AllOE <- 
  leaf_apply(
    models, 
    class = "cesModel", 
    mod2oedf
  ) %>% 
  bind_rows() %>%
  # None of "models" adheres to the CSP.
  mutate (CSP = "w/o CSP")

# Now append to AllOE time-varying output elasticities from models that adhere to the CSP.
# These output elasticities are stored with variable names oe1, oe2, and oe3 in the data object in the models.
# See the <<Models_CSP>> chunk above.
# Uncomment when we're ready to add the CSP models to the alpha vs. time graph.

# AllOE <- leaf_apply(modelsCSP, class="lm", f=function(mod, id){
#   id_components <- strsplit(id, split = "[.]")[[1]]
#   if (id_components[4] == "CES") {
#     return(mod$data)
#   } else {
#     return(NULL)
#   }
# }) %>% 
#   do.call(plyr::rbind.fill, .) %>%
#   bind_rows(AllOE, .)

@

@@

temp_data_CSP <- 
  leaf_apply(modelsCSP, class=c("plm", "lm"), model2df) %>% 
  bind_rows() %>%
  mutate(CSP = "w/ CSP")
